Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-14 talk/pepm2011/paper.tex
Two more comments
fix another an XXX and a typo

The trace shows the inefficiencies of \lstinline{f} clearly, if one looks at the
number of \lstinline{new}, \lstinline{set/get} and \lstinline{guard_class} operations.
In the rest of the paper we will see how this trace can be optimized using
partial evaluation.

XXX add comment that guards are not only slow, but also use lots of memory

@@

The partial evaluation works by traversing the trace from beginning to end.
Whenever a \lstinline{new} operation is seen, the operation is removed and a \empth{static
object}\footnote{Here ``static'' is meant in the sense of partial
evaluation, \ie known at partial evaluation time, not in the sense of static
allocation or static method.} is constructed and associated with the variable
that would have stored
the result of \lstinline{new}. The static object describes the shape of the
original object, \eg where the values that would be stored in the fields of the
allocated object come from, as well as the type of the object. Whenever the
optimizer sees a \lstinline{set} that writes into such an object, that shape
description is updated and the operation can be removed, which means that the
operation was done at partial evaluation time. When the optimizer encounters a
\lstinline{get} from such an object, the result is read from the shape
description, and the operation is also removed. Equivalently, a
\lstinline{guard_class} on a variable that has a shape description can be removed
as well, because the shape description stores the type and thus the outcome of
the type check the guard does is statically known.

@@

@@

>>>>

The trace shows the inefficiencies of \lstinline{f} clearly, if one looks at
the number of \lstinline{new}, \lstinline{set/get} and \lstinline{guard_class}
operations. The number of \lstinline{guard_class} operation is particularly
problematic, not only because of the time it takes to run them. All guards also
have additional information attached that makes it possible to return to the
interpreter, should the guard fail. This means that many guard operations also
lead to a memory problem.

In the rest of the paper we will see how this trace can be optimized using
partial evaluation.

@@

The partial evaluation works by traversing the trace from beginning to end.
Whenever a \lstinline{new} operation is seen, the operation is removed and a \emph{static
object}\footnote{Here ``static'' is meant in the sense of partial
evaluation, \ie known at partial evaluation time, not in the sense of static
allocation or static method.} is constructed and associated with the variable
that would have stored
the result of \lstinline{new}. The static object describes the shape of the
original object, \eg where the values that would be stored in the fields of the
allocated object come from, as well as the type of the object. Whenever the
optimizer sees a \lstinline{set} that writes into such an object, that shape
description is updated and the operation can be removed, which means that the
operation was done at partial evaluation time. When the optimizer encounters a
\lstinline{get} from such an object, the result is read from the shape
description, and the operation is also removed. Equivalently, a
\lstinline{guard_class} on a variable that has a shape description can be removed
as well, because the shape description stores the type and thus the outcome of
the type check the guard does is statically known.

@@

XXX [fijal] Well, in fact that's *a lot* of support for debugger from the VM.
The support is just generic instead of specific.

@@

XXX [fijal] note that some of those have cpython-specific hacks removed or
old versions used (nbody mostly I think).
